Declarative runtime reconfiguration of pulverised distributed systems \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-12 paper-2023-acsos-pulverization.tex
RW: + AASL autonomic spec lang
RW: + ConfSolve + ADSL

\subsection{Automatic deployment and reconfiguration}
\label{sec:rw:autodep}
%
A comprehensive conceptual framework
 on automatic deployment of distributed systems
 is provided in a 2015 survey by Arcangeli et al. \cite{DBLP:journals/jss/ArcangeliBL15}.
%
A more recent survey by Coullon et al. focusses
 on formal techniques for verifying the correctness of reconfigurations
 in component-based distributed software systems~\cite{coullon2023swreconfig}.
%
Two formal models for specifying reconfigurable architectures are \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}.
%
Both are based on the same conceptual model: they 
 feature \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
\emph{maps} (logical topologies),
and \emph{deployments} (associating components to map locations),
overall organised in \emph{motifs} (dynamic architectural configurations),
to model and analyse dynamic architectures.
%
\meta{TODO: say something about our positioning wrt verification of reconfigurations}

In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},
 microservices (the solvent)
 can migrate across 
 the edge-cloud infrastructure (the solution),
 passing through layer boundaries (the semi-permeable membranes),
 in order to keep a balance in the desired properties (the solute).

>>>>

\subsection{Automatic/autonomic deployment and reconfiguration}
\label{sec:rw:autodep}
%
A comprehensive conceptual framework
 on automatic deployment of distributed systems
 is provided in a 2015 survey by Arcangeli et al. \cite{DBLP:journals/jss/ArcangeliBL15}.
%
A more recent survey by Coullon et al. focusses
 on formal techniques for verifying the correctness of reconfigurations
 in component-based distributed software systems~\cite{coullon2023swreconfig}.
%
Two formal models for specifying reconfigurable architectures are \emph{DR-BIP (Dynamically Reconfigurable - Behaviour Interaction Priority model)}~\cite{,Ballouli18dr-bip}
and \emph{DReAM (Dynamically Reconfigurable Architectural Modelling)}~\cite{denicola2020dream-dynamic-reconfig-arch-modelling}.
%
Both are based on the same conceptual model: they 
 feature \emph{components} (capturing behaviour), \emph{connectors} (capturing interaction between components' \emph{ports}), 
\emph{maps} (logical topologies),
and \emph{deployments} (associating components to map locations),
overall organised in \emph{motifs} (dynamic architectural configurations),
to model and analyse dynamic architectures.
%
\meta{TODO: say something about our positioning wrt verification of reconfigurations}

In the \emph{osmotic computing} approach~\cite{DBLP:journals/computer/VillariFDRJR19},
 microservices (the solvent)
 can migrate across 
 the edge-cloud infrastructure (the solution),
 passing through layer boundaries (the semi-permeable membranes),
 in order to keep a balance in the desired properties (the solute).
%
The \emph{ASSL (Autonomic System Specification Language)}~\cite{DBLP:conf/birthday/VassevH97}
 supports the specification
 of autonomic elements,
 interaction protocols,
 and \emph{self-management policies} based on fluents (states), events, and actions.
%
In \emph{ConfSolve}~\cite{DBLP:conf/saso/HewsonAG13},
 the language supports the specification of logical \emph{reconfiguration constraints}, which are then solved as a constraint satisfaction problem.
%
The \emph{AWaRE DSL (ADSL)}~\cite{DBLP:conf/aswec/ChhetriLUVKNR18/adsl} also supports constraint-based self-management, leveraging managing agents. The language enables the specification of the domain model (in terms of components), the problem structure model (in terms of constraints), the agent architecture model (in terms of agents, roles, and their coordination), and the assignment model (in terms of management problem assignment strategies).
